From Reversible Programs to \\ Univalent Universes and Back

### 2021-06-29 Pi+/paper/discussion.tex
More in discussion

\paragraph{Free Symmetric Monoidal Category} The forgetful functor from $\SymMonCat$, the 2-category of symmetric
monoidal categories, strong symmetric monoidal functors, and symmetric monoidal natural transformations, to the
2-category $\CatCat$, has a left adjoint.

Curry-Howard-Lambek correspondence.

>>>>

\paragraph{Free Symmetric Monoidal Category} The forgetful functor from $\SymMonCat$, the 2-category of (small)
symmetric monoidal categories, strong symmetric monoidal functors, and symmetric monoidal natural transformations, to
the 2-category $\CatCat$, has a left adjoint giving the free symmetric monoidal category $\FSM[\CCat{C}]$ on a category
$\CCat{C}$. This is a 2-monad on $\CatCat$~\cite{blackwellTwodimensionalMonadTheory1989}, and its construction is known
in the literature~\cite{abramskyAbstractScalarsLoops2005}. Concretely, the objects of $\FSM[\CCat{C}]$ are given by
lists of objects of $\CCat{C}$, that is, a pair $(n:\Nat, A:[n] \to \CCat{C}_{0})$. An morphism between $(n,A)$ and
$(n,B)$ is given by a pair $(\pi,\lambda)$ where $\pi$ is a permutation of $[n]$, and
$\lambda_{i} : A_{i} \to B_{\pi(i)}$ for $1 \leq i \leq n$. Abstractly, this is given by the Grothendieck construction
$\int F$ of the functor $F : \BFin \to \CatCat$ from the groupoid of finite sets and bijections to $\CatCat$, assigning
each natural number $n$ to the $n$-power $C^{n}$ of $C$, and each permutation on $[n]$ inducing an endofunctor on
$C^{n}$ by action. $\BFin$ is the free symmetric monoidal category (groupoid) on one generator, $\FSM[\unit]$. The free
symmetric monoidal category has been used to study concurrency~\cite{hylandSymmetricMonoidalSketches2004}, petri
nets~\cite{baezCategoriesNets2021}, combinatorial structures~\cite{fioreCartesianClosedBicategory2008}, quantum
mechanics~\cite{abramskyAbstractScalarsLoops2005}, bicategorical models of (differential) linear
logic~\cite{melliesTemplateGamesDifferential2019}. In HoTT, the free symmetric monoidal groupoid $\FSM[A]$ over a
groupoid $A$ can be given by $\dsum{X:\UFin}{A^{X}}$, or it can be presented as an algebraic 2-theory using 1-HITs.
These HITs and the proof of their universal property have been considered
in~\citet*{piceghelloCoherenceSymmetricMonoidal2019,choudhuryFinitemultisetConstructionHoTT2019}.

\paragraph{Curry-Howard-Lambek correspondence} In~\citet{curryCurryEssaysCombinatory1980}, Lambek extended the
Curry-Howard correspondence to cartesian-closed categories. In this work, we have established a correspondence between a
fragment of the $\PiLang$ family of languages and symmetric monoidal groupoids.
